---
title: "Wellcome Global Monitor 2018"
subtitle: "Measuring Knowledge of Science"
format: 
  dashboard: 
    orientation: columns
    nav-buttons: [github, linkedin]
    github: https://github.com/nasoares/lab-3
    linkedin: https://www.linkedin.com/in/nicholas-soares-795a23290/
    theme: sandstone
logo: images/wellcome_logo.png
editor: visual
---

```{r}
#| label: load-packages
#| message: false
library(tidyverse)
library(gt)
library(gtExtras)
library(readxl)
library(ggridges)
library(RColorBrewer)
library(leaflet)
library(sf)
library(viridisLite)
library(gganimate)
library(scales)
library(fontawesome)
```

```{r}
#| label: prep-data
#| message: false
#| warning: false

#Loading spreadsheets
wgm<-read_xlsx("data/wgm2018-dataset-crosstabs-all-countries.xlsx",
               sheet="Full dataset")
notes<-read_xlsx("data/wgm2018-dataset-crosstabs-all-countries.xlsx",
                 sheet="Data dictionary")

#creating a new, filtered data frame
wgm_filt<-wgm|>
  select(c(WP5,wgt,PROJWT,FIELD_DATE,YEAR_CALENDAR,
           Q1,AgeCategories,Regions_Report))

#Handling the country codes
countries<-t(str_split(notes[1,3],',',simplify = T))
countryCode<-str_split(countries,"=",simplify=T)
countryCode<-as.data.frame(countryCode)
colnames(countryCode)<-c("WP5","Country")
countryCode$WP5<-as.numeric(countryCode$WP5)

#Handling the region codes
regions<-t(str_split(notes[57,3],',',simplify=T))
regionCode<-str_split(regions,"=",simplify=T)
regionCode<-as.data.frame(regionCode)
colnames(regionCode)<-c("Regions_Report","Region")
regionCode$Regions_Report<-as.numeric(regionCode$Regions_Report)

#Merging country and region names
wgm_merge<-left_join(wgm_filt,countryCode)
wgm_merge<-left_join(wgm_merge,regionCode)

wgm_country2<-wgm_merge|>
  group_by(Country)|>
  summarize(Total=n())#total records from each country

### Preparing Leaflet Map Data ###
wgm_sci<-wgm_merge|>
  mutate(SciKnow=case_when(
    Q1 <= 2 ~ "Lots",
    Q1 >= 3 ~ "NotLots"
  ))

#Grouping by country for plotting
wgm_sci_country<-wgm_sci|>
  group_by(Country,SciKnow)|>
  summarize(Count=n())#grouping by scientific knowledge
wgm_sci_country<-left_join(wgm_sci_country,wgm_country2)

#Calculation of proportions
wgm_sci_props<-wgm_sci_country|>
  mutate(Prop=Count/Total*100)|>
  filter(SciKnow=="Lots")|>
  select(Country,Prop)
colnames(wgm_sci_props)<-c("NAME","KnowLevel")

### Preparing Animated Plot Data ###
wgm_sci_region<-wgm_sci|>
  filter(!is.na(AgeCategories))|>
  group_by(Region,AgeCategories,SciKnow)|>
  summarize(Count=n())#grouping by scientific knowledge
wgm_sci_region2<-wgm_sci|>
  filter(!is.na(AgeCategories))|>
  group_by(Region,AgeCategories)|>
  summarize(Total=n())
wgm_sci_region<-left_join(wgm_sci_region,wgm_sci_region2)

#Calculation of proportions by age range and region
wgm_sci_age<-wgm_sci_region|>
  mutate(Prop=Count/Total)|>
  filter(SciKnow=="Lots"&Region!="Not assigned")|>
  select(Region,AgeCategories,Prop)

#Calculating World totals
world_age_totals<-wgm_sci_age|>
  group_by(AgeCategories)|>
  summarize(Prop=mean(Prop))|>
  mutate(Region="World")|>
  relocate(Region,.before=AgeCategories)
wgm_sci_age<-rbind(wgm_sci_age,world_age_totals)

#Factoring Region and Age Range for plotting order
wgm_sci_age$Region<-factor(wgm_sci_age$Region,
                           levels = c("World","Eastern Africa","Central Africa",
                                      "North Africa","Southern Africa","Western Africa",
                                      "Central America and Mexico","Northern America",
                                      "South America","Central Asia","East Asia",
                                      "Southeast Asia","South Asia","Middle East",
                                      "Eastern Europe","Northern Europe","Southern Europe",
                                      "Western Europe","Aus/NZ"))
wgm_sci_age<-wgm_sci_age|>
  mutate(AgeRange=case_when(
    AgeCategories == 1 ~ "15 to 29",
    AgeCategories == 2 ~ "30 to 49",
    AgeCategories == 3 ~ "50+"
  ))
wgm_sci_age$AgeRange<-factor(wgm_sci_age$AgeRange,
                             levels = c("50+","30 to 49","15 to 29"))
```

# Results by Country

## Column - Leaflet Map {width="70%"}

```{r}
#| label: leaflet-map
#| title: "Perceived Knowledge About Science By Country"
#| message: false
#| warning: false

#Loading in shapefile of countries for chloropleth map
#following guidelines outlined in https://r-graph-gallery.com/183-choropleth-map-with-leaflet.html
world_sf<-read_sf("data/TM_WORLD_BORDERS_SIMPL-0.3.shp")

wgm_sf<-world_sf|>
  filter(NAME %in% unique(wgm_sci_props$NAME))
wgm_sf<-left_join(wgm_sf,wgm_sci_props)

#Creating a color palette to display data
mypalette<-colorNumeric(
  palette="viridis",domain=wgm_sf$KnowLevel,
  na.color="transparent"
)

#Text labels
mytext<-paste(
  "Country: ",wgm_sf$NAME,"<br/>",
  "Knowledge Level: ",round(wgm_sf$KnowLevel,0),"%","</br>",
  sep=""
)%>%
  lapply(htmltools::HTML)

#Display map
leaflet(wgm_sf)%>%
  addTiles()%>%
  setView(lat=0,lng=0,zoom=2)%>%
  addPolygons(fillColor = ~mypalette(KnowLevel),stroke = F,
              label=mytext,fillOpacity = 0.9)%>%
  addLegend(pal = mypalette,values=~KnowLevel,
            title="Self-Reported </br> Knowledge Level (%)",position = "bottomleft",
            opacity=0.9,labFormat = labelFormat(suffix = "%"))
```

## Column - Data Cards {width="30%"}

```{r}
#| label: quick-stats

knowlevel_high<-wgm_sf|>
  select(NAME,KnowLevel)|>
  slice_max(order_by = KnowLevel)

knowlevel_low<-wgm_sf|>
  select(NAME,KnowLevel)|>
  slice_min(order_by = KnowLevel)
```

::: {.valuebox icon="globe2" color="#5DC863"}
Average of Surveyed Countries:

`{r} round(mean(wgm_sf$KnowLevel),1)`
:::

::: {.valuebox icon="globe-americas" color="#FDE725"}
Highest Knowledge of Science:

`{r} round(knowlevel_high$KnowLevel,0)`

`{r} knowlevel_high$NAME`
:::

::: {.valuebox icon="globe-europe-africa" color="#21908C"}
Lowest Knowledge of Science:

`{r} round(knowlevel_low$KnowLevel,0)`

`{r} knowlevel_low$NAME`
:::

# Results by Region and Age

## Column - animated plot {width="60%"}

```{r}
#| label: animate-map
#| title: "Perceived Knowledge About Science By Region and Age Group"
#| message: false
#| warning: false
#| fig-width: 15
#| fig-height: 10

#Plotting with gganimate
ggplot(wgm_sci_age,aes(Prop,Region,fill=AgeRange),color="black")+
  geom_bar(stat="identity",position = "dodge")+
  scale_y_discrete(limits=rev)+
  scale_fill_manual(values=c("#003366","#33CCCC","#FFCC00"),
                    guide=guide_legend(reverse = T))+ #liked the color scheme of the original plot
  scale_x_continuous(labels = scales::percent_format(accuracy = 1))+
  labs(x=NULL,y=NULL,title=NULL)+
  theme_minimal()+
  theme(legend.position = c(0.85,0.8),
        panel.grid.minor.x = element_blank(),
        panel.grid.major.y = element_blank(),
        text = element_text(size=20),
        panel.background = element_rect(fill = "#faf9f6",color = "#faf9f6"),
        plot.background = element_rect(fill="#faf9f6",color="#faf9f6"),
        panel.grid.major.x = element_line(linetype = "dashed",colour = "grey"),
        plot.title.position = "plot")+
  transition_states(Region,wrap=F)+
  shadow_mark()+
  enter_fade()
```

## Overview

### Summary

All data displayed on this dashboard comes from Chapter 2 of the [2018 Wellcome Global Monitor Report](https://cms.wellcome.org/sites/default/files/wellcome-global-monitor-2018.pdf), entitled "Understanding and interest in science and health".

The primary metric used in this chapter of the report is "Knowledge Level", defined as the percentage of people who answered 'a lot' or 'some' in response to the question: *How much do you know about science? Do you know a lot, some, not much, or nothing at all?*

### Tabset {.tabset width="40%"}

::: {.card title="Age Summary"}
Age ranges sorted in descending order of knowledge level (%). In general, a higher percentage of 15-29 year olds across all regions self-report having 'a lot' or 'some' scientific knowledge compared to the older age groups.

```{r}
#| label: age-table

age_table<-wgm_sci_age|>
  group_by(AgeRange)|>
  summarize(Prop=mean(Prop))|>
  arrange(desc(Prop))|>
  mutate(Percentage=paste(round(Prop,2)*100,"%",sep=""))|>
  select(AgeRange,Percentage)|>
  gt()|>  
  cols_align(align="left",
             columns = c("AgeRange","Percentage"))|>
  gt_theme_guardian()
age_table
```
:::

::: {.card title="Region Summary"}
Regions sorted in descending order of knowledge level (%). The list of countries belonging to each region have been defined on page

```{r}
#| label: region-table

region_table<-wgm_sci_age|>
  group_by(Region)|>
  summarize(Prop=mean(Prop))|>
  arrange(desc(Prop))|>
  mutate(Percentage=paste(round(Prop,2)*100,"%",sep=""))|>
  select(Region,Percentage)|>
  gt()|>
  cols_align(align="left",
             columns = c("Region","Percentage"))|>
  gt_theme_guardian()
region_table
```
:::
